---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(here)
library(tidyverse)
load(here("test.RData"))
```

# LifelinesDataDashboard
>Date: 5-01-2025

<!-- badges: start -->
<!-- badges: end -->

The goal of the package LifelinesDataDashboard is to visualize the Lifelines dataset and to show the significance of correlations between gender, age, income and many more. 
This dashboard has interactive functionality to make visualizations based on filters of you own choosing, like age and gender. 

## Installation

You can install the development version of LifelinesDataDashboard from [GitHub](https://github.com/) with:

``` {r, eval=F}
# First install het pak library:
install.packages("pak")

# Then, use pak to install the package from github:
pak::pak("Fl-ris/Lifelines_data_dashboard")
```



## Example
The code snippets below demonstrate some of the functionality of the LifelinesDataDashboard package.

First, load the library.
```{r run-datadashboard, eval=T}
library(LifelinesDataDashboard)
```

After that, load the dataset with the function: `run_datadashboard()` and give the path the the dataset as shown in the following example:
```{r, eval=F}

#run_datadashboard(here("Data", "Lifelines Public Health dataset - 2024.csv"))

load_dataset(here("Data", "Lifelines Public Health dataset - 2024.csv"))
```


The plotting functions can also be called upon without the shiny interface:
```{r Gender_distribution_paticipants, eval=T}

gender_dist(dataset = lifelines_df)
```


```{r Participant-weight-distribution, eval = T}
weight_dist(dataset = lifelines_df)
```

```{r Wealth_and_satisfaction_correlation, eval = T}
finance_neighborhood_cor(dataset = lifelines_df)
```

**Dependencies:** \
Dependencies for the LifelinesDataDashboard library: \
- R (>= 3.5.0) \
- pak \
- Other libraries will be automatically installed with 'pak("Fl-ris/Lifelines_data_dashboard")' \

Dependencies for the Notebook with my analysis: \
- R \
- tidyverse \
- stringr \


**Author:**
F.J.A. Menninga, github name: fl-ris




