---
title: "Lifelines Datadashboard"
output: html_notebook
---


### Introduction ###



**Data:**

The data used is a synthetic dataset, this means that the values are not the exact same as the ones that were measured but a statisticallcly representative dataset. 
This is done to enhance privacy of the participants. The methods used for this are: differential privacy and k-anonymity. This way the relationsships between the variables are conserverd. 
Source: https://www.lifelines-biobank.com/synthetic-data-a-new-step-forward-in-data-availability-at-lifelines-in-collaboration-with-syntho



```{r}

library(tidyverse)

```



```{r}
lifelines_df <- read.csv(file = "/home/floris/Documenten/Data_set/Lifelines/2024/Dataset/Lifelines Public Health dataset - 2024.csv", header = TRUE)

```


```{r}

lifelines_long <- lifelines_df %>% pivot_longer(cols = GENDER)
  

ggplot(data = lifelines_long, mapping = aes(x = HEIGHT_T1)) +
  geom_density() +
  xlab("") +
  ylab("") +
  theme_minimal()

  

```


```{r}

neighborhood_satisfaction <- lifelines_df$NEIGHBOURHOOD1_T2 %>%  factor(levels = c(1:10), labels = c(1:10))

neighborhood_characteristics <- lifelines_df$NEIGHBOURHOOD2_T2 %>%  factor(levels = c(1:10), labels = c(1:10))

neighborhood_Unpleasantness <- lifelines_df$NEIGHBOURHOOD3_T2 %>%  factor(levels = c(1:10), labels = c(1:10))

neighborhood_Unpleasantness <- lifelines_df$NEIGHBOURHOOD4_T2 %>%  factor(levels = c(1:10), labels = c(1:10))

```



```{r}



```




