---
title: "Lifelines Datadashboard"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


### Introduction ###



**Data:**

The data used is a synthetic dataset, this means that the values are not the exact same as the ones that were measured but a statisticallcly representative dataset. 
This is done to enhance privacy of the participants. The methods used for this are: differential privacy and k-anonymity. This way the relationsships between the variables are conserverd. 
Source: https://www.lifelines-biobank.com/synthetic-data-a-new-step-forward-in-data-availability-at-lifelines-in-collaboration-with-syntho


**Libraries**
The following libraries are used:
```{r}

library(tidyverse)
library(stringr)
```



```{r}
# Without factors (test)
lifelines <- read.csv(file = "/home/floris/Documenten/Data_set/Lifelines/2024/Dataset/Lifelines Public Health dataset - 2024.csv", header = TRUE)

lifelines_df <- read.csv(file = "/home/floris/Documenten/Data_set/Lifelines/2024/Dataset/Lifelines Public Health dataset - 2024.csv", header = TRUE)

```



### Making factors ###
Making factors for the schale data using the following variables:

NEIGHBOURHOOD1_T2 'Satisfaction with the current living environment (scored on a scale of 1 to 10) at second assessment (T2)'

NEIGHBOURHOOD2_T2 'Characteristics of neighbourhood (ranging from a very green neighbourhood (1) to a neighbourhood with practically no greenery (5)

NEIGHBOURHOOD3_T2 'Unpleasantness to live in this neighbourhood (ranging from completely disagree (1) to completely agree (5)

NEIGHBOURHOOD4_T2 'If possible, would like to move from this neighbourhood (ranging from completely disagree (1) to completely agree (5))

NEIGHBOURHOOD5_T2 'Attached to this neighbourhood (ranging from completely disagree (1) to completely agree (5))

NEIGHBOURHOOD6_T2 'Feels at home in this neighbourhood (ranging from completely disagree (1) to completely agree (5))


The following factors are about socioeonomically determined variables:

```{r Socioeconomical-factors}

lifelines_df$neighborhood_satisfaction <- lifelines_df$NEIGHBOURHOOD1_T2 %>%  factor(levels = c(1:10), labels = c(1:10))

lifelines_df$neighborhood_characteristics <- lifelines_df$NEIGHBOURHOOD2_T2 %>%  factor(levels = c(1:5), labels = c(1:5))

lifelines_df$neighborhood_unpleasantness <- lifelines_df$NEIGHBOURHOOD3_T2 %>%  factor(levels = c(1:5), labels = c("Completely agree", "Agree", "Neutral", "Disagree", "Completely disagree"))

lifelines_df$neighborhood_moving_away <- lifelines_df$NEIGHBOURHOOD4_T2 %>%  factor(levels = c(1:5), labels = c("Completely agree", "Agree", "Neutral", "Disagree", "Completely disagree"))

lifelines_df$neighborhood_attached <- lifelines_df$NEIGHBOURHOOD5_T2 %>%  factor(levels = c(1:5), labels = c("Completely agree", "Agree", "Neutral", "Disagree", "Completely disagree"))

lifelines_df$neighborhood_feel_at_home <- lifelines_df$NEIGHBOURHOOD6_T2 %>%  factor(levels = c(1:5), labels = c("Completely agree", "Agree", "Neutral", "Disagree", "Completely disagree"))

lifelines_df$FINANCE_T1 <- lifelines_df$FINANCE_T1 %>% factor(levels = c(1:10),  labels = c("I do not know", "I don't want to answer", "Less than 750", "750 - 1000", "1000 - 1500", "1500 - 2000", "2000 - 2500", "2500 - 3000", "3000 - 3500", "More than 3500"))

# Quality of life, 0 = low QOL and 1 = 
lifelines_df$LOW_QUALITY_OF_LIFE_T1  <- lifelines_df$LOW_QUALITY_OF_LIFE_T1 %>%  factor(levels = c(0:1), labels = c("Mediocre", "False"))

lifelines_df$LOW_QUALITY_OF_LIFE_T2 <- lifelines_df$LOW_QUALITY_OF_LIFE_T2 %>%  factor(levels = c(0,1), labels = c("Low/Mediocre", "False"))


lifelines_df$SMOKING <- lifelines_df$SMOKING %>%  factor(levels = c(0,1), labels = c("Smoking", "Non-smoking"))

```

The following factors are about body "specifications" like age and weight.


```{r Personal-factors}


lifelines_df$GENDER <- lifelines_df$GENDER %>%  factor(levels = c(1:2), labels = c("Male", "Female"))


```


```{r}



```

The following figure displays the distribution of height:
```{r density-plot-age}

# To make the data in the long format:
#lifelines_long <- lifelines_df %>% pivot_longer(cols = ZIP_CODE)
  
# Plot it using facet_wrap to see the difference between the genders:
ggplot(data = lifelines_df, mapping = aes(y = HEIGHT_T1)) +
  geom_boxplot(fill = "darkblue", alpha = 0.7) +
  xlab("") +
  ylab("Height") +
  facet_wrap(~GENDER) +
  ggtitle("Height distribution") +
  theme_minimal()



```

```{r}

#ggplot(data = lifelines_df, mapping = aes(y =  ))

```

21-11-2024:

Using facet_wrap, the differences in weight distribution is displayed in de following plot.
Even though the mean weight for the females is lower than the one of the males, the spread is higher, this was quantified using ....




```{r weight-distribution}

ggplot(data = lifelines_df, mapping = aes(y = WEIGHT_T1) ) +
  geom_boxplot(fill = "green", alpha = 0.5) +
  xlab("Count: ") +
  ylab("Weight") +
  ggtitle("Participant weight:") +
  facet_wrap(~GENDER) +
  theme_minimal()


```

For the woman, the amount of pregnacies was visualized in the following plot:

```{r}

ggplot(data = lifelines_df, mapping = aes(y = PREGNANCIES)) +
  geom_boxplot(fill = "blue", alpha=0.6) +
  xlab("") +
  ylab("Pregnancies:") +
  theme_minimal()
```


```{r gender-distribution, eval=F}
ggplot(data = lifelines_df, mapping = aes(y = GENDER) ) +
  geom_bar(fill = "blue", alpha = 0.6) +
  xlab("Count: ") +
  ylab("Gender") +
  ggtitle("Participant count:") +
  theme_minimal()


```
```{r}

FINANCE <- lifelines_df[!is.na(lifelines_df$FINANCE_T1), ]
#FINANCE <- subset(lifelines_df[FINANCE])

ggplot(data = lifelines_df, mapping = aes(x = neighborhood_satisfaction, y = FINANCE_T1)) +
  geom_jitter(mapping = aes(alpha = 0.5)) +
  xlab("") +
  ylab("") +
  ggtitle("Correlation between wealth and neighborhood satisfaction") +
  theme_minimal()


```
todo heatmap!
The plot above displays the neighborhood satisfaction score for every salary class. 
There is a clear trend where people who earn more are more satisfied with their neighborhood. 


### Using Generalized linear models to quantify correlations: ###
A generalized linear model is used to see if there is a correlation between for example; the amount of money people make monthly and what their opinion is on their living arrangements, mainly the neighborhood in wich they live. 
In this codechunk, this opinion is used as outcome variable and montly income is the predictor variable. As the summary states, there is a significant correlation between these two variable.
The p-value is: 0.000198 *** for 

Here the variables used are opinion of living  arragements as outcome variable and education as predictor. There appears to be a sigificant correlation between these two variables. 


```{r}
neighborhood_prediction <- glm(formula = lifelines_df$NEIGHBOURHOOD1_T2 ~ lifelines_df$EDUCATION_LOWER_T1 + lifelines_df$NEIGHBOURHOOD4_T2, na.action = na.exclude)

neighborhood_prediction_p <- predict(neighborhood_prediction)

summary(neighborhood_prediction)
```
According to the glm() prediction above, both education and the wish to move away from their current house are significant predictors for the satisfaction score.  
The latter predictors is obvious, if people want to move away, they are most likely not satisfied with their current place. 

```{r}
plot(neighborhood_prediction)

```
27-11-2024:


#```{r}
# Removing the NA values from the data that will be used for the logistic regression:


#```

```{r eval=F}
neighborhood <- glm(formula = SMOKING ~ EDUCATION_LOWER_T1, data = lifelines)

predicted <- predict(neighborhood)

summary(neighborhood_prediction)

plot(test2)
```

```{r test1, eval=F}
# Fit the glm model using 'na.exclude'
neighborhood <- glm(formula = SMOKING ~ EDUCATION_LOWER_T1, data = lifelines_df, na.action = na.exclude)

# Get the fitted values and residuals
fitted_values <- neighborhood$fitted.values

# Identify the rows used in fitting the model
used_indices <- !is.na(neighborhood$residuals)  # This identifies which rows were actually used

# Subset both fitted values and smoking data to match these indices
smoking_subset <- lifelines$SMOKING[rownames(lifelines)[used_indices]]

# Now create your dataframe with matching lengths
new_data <- data.frame(fitted_values, smoking_subset)

# Optionally give the columns meaningful names
colnames(new_data) <- c("Fitted_SMOKING", "Actual_SMOKING")
```

### Statistical tests:

First, to check for a normal distribution, I wanted to use the Shapiro-wilk test on the data. 
The Shapiro-wilk test can only handle up to 5000 observations and given that this Lifelines dataset contains over 16000, this test can not be used. 
Therefore, the anderson-darling test will be used. 

```{r, eval=F}
library(nortest)

normality_height <- ad.test(lifelines_df$HEIGHT_T1)
normality_height



```
```{r}

test3 <- glm(formula = lifelines$RESPIRATORY_DISEASE_T1 ~ lifelines_df$SMOKING + lifelines_df$LOW_QUALITY_OF_LIFE_T1, na.action = na.exclude)

test2_p <- predict(test3)

summary(test3)

plot(test3)




```
28-11-2024:

```{r, eval=F}

lifelines_male <- lifelines_df %>% 
  filter(GENDER == "Male")

lifelines_female <- lifelines_df %>% 
  filter(GENDER == "Female")

#filter_function <- function(option) {
#  output <- lifelines_df %>% 
#    filter(GENDER == fpaste(option))
#  
#}

ad_mheight <- ad.test(lifelines_male$HEIGHT_T1)

ad_fheight <- ad.test(lifelines_female$HEIGHT_T1)

ad_mweight <- ad.test(lifelines_male$WEIGHT_T1)

ad_fweight <- ad.test(lifelines_female$WEIGHT_T1)

ad_mbmi <- ad.test(lifelines_male$BMI_T1)
  
ad_fbmi <- ad.test(lifelines_female$BMI_T1)

p <- c(ad_mheight$p.value, ad_fheight$p.value)



```

```{r, eval=F}

t.test(lifelines_female$HEIGHT_T1, lifelines_male$HEIGHT_T1)

```
4-12-2024:

PCA:
Below is a principal component analysis to visualize the many dimensions in one 2d plane. 
The following code does not work yet. 
```{r, eval=F}

pr <- princomp(t(lifelines_df))

biplot(pr)

d16 <- dist(t(rnaseq[1:6,3:18]))

image(as.matrix(d16))

```

```{r eval=F}
test <- lapply(lifelines_df$AGE_T1, FUN = sum(n)


plot3d( lifelines_df[,1], lifelines_df[,2], lifelines_df[,3], col = lifelines_df$HEIGHT_T1, type = "s", radius = .2 )

play3d( spin3d( axis = c(0, 0, 1), rpm = 20), duration = 10 )

```

Because of the normally distributed data, the t-test can be used to measure a difference between the two groups. 
```{r eval=F}

height_men <- lifelines_df %>% 
  filter(GENDER == "Male")

height_women <- lifelines_df %>% 
 filter(GENDER == "Female")


#T-test:
t.test(x = height_men, y = height_women, alternative = "two.sided")



```


```{r eval=F}
# Differences between:
# Men and woman:

# Smoking:
Smoking_MW <- t.test(as.numeric(lifelines_female$SMOKING), as.numeric(lifelines_male$SMOKING))

# Education:
Education_MW <- t.test(as.numeric(lifelines_female$EDUCATION_LOWER_T1), as.numeric(lifelines_male$EDUCATION_LOWER_T1))

# Height:
Height_MW <- t.test(as.numeric(lifelines_female$HEIGHT_T1), as.numeric(lifelines_male$HEIGHT_T1))

# Respiratory disease:
Resp_dis_MW <- t.test(as.numeric(lifelines_female$RESPIRATORY_DISEASE_T1), as.numeric(lifelines_male$RESPIRATORY_DISEASE_T1))

# BMI:
Bmi_MW <- t.test((lifelines_female$BMI_T1), (lifelines_male$BMI_T1))

# Low QOL:
qol_MW <- t.test(as.numeric(lifelines_female$LOW_QUALITY_OF_LIFE_T1), as.numeric(lifelines_male$LOW_QUALITY_OF_LIFE_T1))

# Waist size:
Waist_MW <- t.test(as.numeric(lifelines_female$WAIST_T1), as.numeric(lifelines_male$WAIST_T1))






```



```{r eval=F}
wilcox_test_vector_geslacht <- c(Smoking_MW[3],Education_MW[3],Height_MW[3],Resp_dis_MW[3],Bmi_MW[3],qol_MW[3],Waist_MW[3])

df_wilcox_geslacht <- (data.frame((wilcox_test_vector_geslacht)) %>% rename(
  `Smoking` = p.value,
  `Education` = p.value.1,
  `Height` = p.value.2,
  `Respitoroty dis` = p.value.3,
  `BMI` = p.value.4,
  `Quality of life` = p.value.5,
  `Waist size` = p.value.6
  ))

df_wilcox_geslacht %>% 
  pivot_longer(cols = c(1:7)) -> df_wilcox_geslacht
```



```{r Visualization-of-P-values-of-T-tests, eval=F}

ggplot(data = df_wilcox_geslacht, mapping = aes(x = name , y = value)) +
  geom_bar(stat = "identity",, fill = "blue", alpha = 0.5) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  geom_label_repel(position = position_nudge_repel(y = -0.07),aes(label = round(value,digits = 3)) ) +
  ylab("P-value:") +
  xlab(" ") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.6)) 
  

```

18-12-2024:










